Visible Minorities and Economic Inequality: An Examination of Rosedale-Moore Park and Downtown Yonge East

### 2024-09-25 outputs/paper/paper.qmd
rendered paper to PDF
adding captions and fixing display issues

format: pdf
number-sections: true
bibliography: references.bib
---

@@

The remainder of this paper is structured as follows: @sec-data presents the methods used to analyze these demographic differences, followed by @sec-results, which details the results of the comparison between Rosedale-Moore Park and Downtown Yonge East. @sec-discussion offers a discussion of the findings, highlighting the implications for understanding residential segregation in Toronto. Finally, @sec-conclusionconcludes with a reflection on the limitations of the study and suggestions for future research directions.

@@

### Data Analysis
To compare the two neighborhoods, I created four bar charts representing different demographic aspects: visible minority groups, income levels, highest education attained, and knowledge of official languages. Each chart displayed the number of individuals on the y-axis, with the x-axis representing the respective categories. The charts were color-coded by neighborhood, facilitating easy visual comparison between Rosedale-Moore Park and Downtown Yonge East. This approach allowed for a clear assessment of the disparities and similarities in demographic profiles without the need for summary statistics.

@@

The dataset is comprised of several key features being compared between the two neighborhoods, including demographic, social and economic characteristics. These features are visible minority, income, highest level of education and knowledge of official language.

### Visible Minority

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-one
#| fig-cap: The bar chart shows the distribution of visible and non-visible minority groups in two of Toronto's neighborhoods.

@@

if (file.exists(file_path)) {
  analysis_data <- read_csv(file_path)
  print("File found and loaded successfully!")
} else {
  stop("Error: File not found at the specified path.")
}

@@

## Graphing visible minority data
# Define categories
visible_minority_categories <- c("South Asian", "Chinese", "Black", "Filipino",
                                 "Arab", "Latin American", "Southeast Asian",
                                 "West Asian", "Korean", "Japanese",
                                 "Visible minority, n.i.e.",
                                 "Multiple visible minorities",
                                 "Not a visible minority")

@@

# Create the plot with number of individuals
visible_minorities_plot <- ggplot(filtered_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Visible Minority Categories Between Neighborhoods",
       x = "Visible Minority Categories",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "lightgreen", "Downtown Yonge East" = "forestgreen")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_data$Value), by = 2000))  # Adjust 'by' value as needed

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-two
#| fig-cap: The bar chart shows the distribution of income in two of Toronto's neighborhoods.

@@

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-three
#| fig-cap: The bar chart shows the distribution of highest level of education attainment in two of Toronto's neighborhoods.

@@

@@

# Create the plot with the number of individuals for education attainment
education_plot <- ggplot(filtered_education_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Education Attainment Between Neighborhoods",
       x = "Education Levels",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "lightblue", "Downtown Yonge East" = "blue")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +  # Rotate x-axis labels to 90 degrees
  scale_y_continuous(breaks = seq(0, max(filtered_education_data$Value), by = 500))  # Adjust 'by' value for more tick marks

# Save the plot with increased vertical size
ggsave("education_plot.png", plot = education_plot, width = 10, height = 20)

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-four
#| fig-cap: The bar chart shows the distribution of visible and non-visible minority groups in two of Toronto's neighborhoods.

@@

# Create the plot with number of individuals for knowledge of official languages
language_plot <- ggplot(filtered_language_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Knowledge of Official Languages Between Neighborhoods",
       x = "Knowledge of Official Languages",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "yellow", "Downtown Yonge East" = "gold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_language_data$Value), by = 2000))  # Adjust 'by' value for more tick marks

@@

# Results

## Overview
In this section, I present the results of the analysis comparing Rosedale-Moore Park and Downtown Yonge East across four key demographic features: visible minority groups, income levels, highest levels of education attained, and knowledge of official languages. The following figures illustrate the distributions and comparisons for each of these features.

## Visible Minority Distribution
Figure 1 shows the distribution of visible minority and non-minority groups in Rosedale-Moore Park and Downtown Yonge East. The grouped bar chart indicates that Rosedale-Moore Park has a significantly higher number of non-visible minorities compared to Downtown Yonge East, with more than double the population in this category. Notably, the South Asian and Chinese communities are the most prominent visible minority groups in both neighborhoods, but Downtown Yonge East has a greater overall presence of visible minorities. This suggests that the downtown area is more culturally diverse, which may influence social dynamics and access to community resources.

@@

# Discussion

@@

The analysis of visible minority distribution between Rosedale-Moore Park and Downtown Yonge East reveals significant demographic contrasts. Rosedale-Moore Park's predominance of non-visible minorities, coupled with a smaller visible minority population, suggests a more homogeneous cultural environment. In contrast, Downtown Yonge East's diversity may contribute to richer cultural dynamics and social interactions, but it could also present challenges in terms of resource allocation and community cohesion. Understanding the implications of these demographic differences is crucial for developing targeted social programs that address the needs of diverse populations.

@@

# Conclusion

@@

# Appendix {-}

>>>>

format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

@@

The remainder of this paper is structured as follows: @sec-data presents the methods used to analyze these demographic differences, followed by @sec-results, which details the results of the comparison between Rosedale-Moore Park and Downtown Yonge East. @sec-discussion offers a discussion of the findings, highlighting the implications for understanding residential segregation in Toronto. Finally, @sec-conclusion concludes with a reflection on the limitations of the study and suggestions for future research directions.

@@

### Data Analysis
To compare the two neighborhoods, I created four bar charts representing different demographic aspects: ethnic groups, income levels, highest education attained, and knowledge of official languages. Each chart displayed the number of individuals on the y-axis, with the x-axis representing the respective categories. The charts were color-coded by neighborhood, facilitating easy visual comparison between Rosedale-Moore Park and Downtown Yonge East. This approach allowed for a clear assessment of the disparities and similarities in demographic profiles without the need for summary statistics.

@@

The dataset is comprised of several key features being compared between the two neighborhoods, including demographic, social and economic characteristics. These features are ethnicity, income, highest level of education and knowledge of official language.

### Ethnicity

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-one
#| fig-cap: The chart illustrates the distribution of various ethnic groups across two Toronto neighborhoods, Rosedale-Moore Park and Downtown Yonge East, highlighting differences in ethnic diversity.

@@

if (file.exists(file_path)) {
  analysis_data <- read_csv(file_path)
  invisible("File found and loaded successfully!")
} else {
  stop("Error: File not found at the specified path.")
}

@@

## Graphing ethnicity data
# Define categories
visible_minority_categories <- c("South Asian", "Chinese", "Black", "Filipino",
                                 "Arab", "Latin American", "Southeast Asian",
                                 "West Asian", "Korean", "Japanese",
                                 "Visible minority, n.i.e.",
                                 "Multiple visible minorities",
                                 "Not a visible minority")

@@

# Create the plot with number of individuals
visible_minorities_plot <- ggplot(filtered_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Ethnic Groups Between Neighborhoods",
       x = "Ethnic Groups",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "lightgreen", "Downtown Yonge East" = "forestgreen")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_data$Value), by = 2000))  # Adjust 'by' value as needed

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-two
#| fig-cap: This bar chart compares the distribution of income categories in Rosedale-Moore Park and Downtown Yonge East, highlighting significant differences in income distribution and reflecting thier economic diversity.

@@

# Filter the data for income-related rows
filtered_income_data <- long_data %>%
  filter(`Neighbourhood Name` %in% income_categories)

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-three
#| fig-cap: This bar chart illustrates the distribution of educational attainment levels in Rosedale-Moore Park and Downtown Yonge East, revealing disparities in educational qualifications and differences in access to education opportunies.

@@

# Filter the data for education-related rows
filtered_education_data <- long_data %>%
  filter(`Neighbourhood Name` %in% education_categories)

@@

# Create the plot with the number of individuals for education attainment
education_plot <- ggplot(filtered_education_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Education Attainment Between Neighborhoods",
       x = "Education Levels",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "lightblue", "Downtown Yonge East" = "blue")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +  # Rotate x-axis labels to 90 degrees
  scale_y_continuous(breaks = seq(0, max(filtered_education_data$Value), by = 500))

# Save the plot with increased vertical size
ggsave("education_plot.png", plot = education_plot, width = 30, height = 49)

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-four
#| fig-cap: This bar chart depicts the distribution of knowledge of official languages among residents of Rosedale-Moore Park and Downtown Yonge East, highlighting variation in language knowledge, reflecting their diverse linguistic profiles.

@@

# Create the plot with number of individuals for knowledge of official languages
language_plot <- ggplot(filtered_language_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Knowledge of Official Languages \n Between Neighborhoods",
       x = "Knowledge of Official Languages",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "yellow", "Downtown Yonge East" = "gold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_language_data$Value), by = 2000))  # Adjust 'by' value for more tick marks

@@

# Results {#sec-results}

## Overview
In this section, I present the results of the analysis comparing Rosedale-Moore Park and Downtown Yonge East across four key demographic features: ethnicity, income levels, highest levels of education attained, and knowledge of official languages. The following figures illustrate the distributions and comparisons for each of these features.

## Ethnicity Distribution
Figure 1 shows the distribution of ethnic groups in Rosedale-Moore Park and Downtown Yonge East. The grouped bar chart indicates that Rosedale-Moore Park has a significantly higher number of non-visible minorities compared to Downtown Yonge East, with more than double the population in this category. Notably, the South Asian and Chinese communities are the most prominent visible minority groups in both neighborhoods, but Downtown Yonge East has a greater overall presence of visible minorities. This suggests that the downtown area is more culturally diverse, which may influence social dynamics and access to community resources.

@@

# Discussion {#sec-discussion}

@@

The analysis of ethnic group distribution between Rosedale-Moore Park and Downtown Yonge East reveals significant demographic contrasts. Rosedale-Moore Park's predominance of non-visible minorities, coupled with a smaller visible minority population, suggests a more homogeneous cultural environment. In contrast, Downtown Yonge East's diversity may contribute to richer cultural dynamics and social interactions, but it could also present challenges in terms of resource allocation and community cohesion. Understanding the implications of these demographic differences is crucial for developing targeted social programs that address the needs of diverse populations.

@@

# Conclusion {#sec-conclusion}

@@

# Appendix {#sec-appendix}
